Measuring individual differences in the understanding of gaze cues

### 2022-08-24 paper/manuscript.Rmd
comments discussion Manuel

---
title             : "Measuring individual differences in the understanding of gaze cues across the lifespan"
shorttitle        : "gaze understanding"
author:

@@

  - name          : "Daniel M. Haun"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"
authornote: |
abstract: |
  To navigate the social world and interact with other agents, we use social cognition. Traditional measures of social cognition often lack satisfactory psychometric properties and are not designed to capture variation *between* individuals. This is problematic because developmental theories emphasize the role of social interaction on the development of social cognition which varies substantially across individuals. To fully understand how social-cognitive abilities emerge and relate to each other, we need new tools that can reliably measure individual differences.
   Here, we present a new task to study individual differences in social cognition in children and adults. We concentrate on an essential ability that is involved in many social-cognitive reasoning processes: gaze understanding – the ability to locate and use the attentional focus of an agent. Our interactive browser-based task works across devices and enables in-person and remote testing. The implemented spatial layout allows for discrete and continuous measures of participants’ click imprecision and is easily adaptable to different study requirements. 
   Our task induces inter-individual differences in a child (N = 387) and an adult (N = 236) sample. Our two study versions and data collection modes yield comparable results that show substantial developmental gains: the older children are, the more accurately they locate the target. High internal consistency and test-retest reliability estimates underline that the captured variation is systematic. This work shows a promising way forward in the study of individual differences in social cognition and will help us explore the in(ter)dependence of our core social-cognitive processes in greater detail.
  All materials and code are freely accessible to interested researchers. The task can be found here: https://ccp-odc.eva.mpg.de/gafo-demo/.

@@

Designing this task required a new testing infrastructure.
We designed the task as an interactive browser-based web app.
This greatly increased the flexibility with which we could modify the stimuli on a trial-by-trial basis.
Furthermore, because the task is largely self-contained, it is much more controlled and standardized.
Most importantly, it makes the task portable: testing is possible in-person using tablets but also remotely via the internet (no installation needed).
As such, it provides a solid basis to study individual differences in gaze understanding across ages at scale.
We make the task and its source code openly accessible for other researchers to use and modify.

@@

@@

Adding complexity to our gaze understanding task might increase individual differences in an adult sample.
Researchers could consider adjusting the task difficulty, for example, by changing the target speed or implementing a saccade version (i.e., where the agent only briefly looks at the target position and then back to the center).
Another task extension could realize differential feedback dependent on the participant's accuracy.
By inducing more between-person variance, reliability estimates could further increase.

As we could see in the split-half reliability calculations, the more accurately the statistical method represents the task structure, the higher our reliability estimates are.
Therefore, we argue that future research should aim at implementing statistical analyses that mirror the experimental design in its complexity.
Computational cognitive models are a promising approach forward.
We could take advantage of all available information and model variation between and within individuals in an even more fine-grained and psychologically interpretable manner.
Computational frameworks could also be used to model performance and their underlying cognitive processes across tasks.
With nested hierarchical models, we could assess the systematic relation between various social-cognitive abilities and recover potentially shared structures between cognitive processes [@bohn2022individual].

In addition to the new task design itself, we designed a new testing infrastructure.
Our gaze understanding task is presented as an interactive web app.
This enables presentation across devices without any prior installation.
Stimuli presentation is achieved through the use of SVGs.
This has several advantages: the aspect ratio and stimulus quality are kept constant no matter which size the web browser displays.
The cartoon-like presentation makes the task engaging for children and adults alike.
Most importantly, we can dynamically modify the stimulus details (e.g., target positions) on a trial-by-trial basis.
Presented agents and objects can be easily adapted for future task modifications or specific cultural settings.

@@

After having probed our new testing infrastructure and psychometric properties, the next step will be to ensure the validity of our gaze understanding task.
Validity is often assessed by looking at concurrent relations between measures.
Studies on gaze following traditionally present children with a simultaneous, congruent movement of eyes and head orientation.
In our study, however, children only observe a subtle eye movement.
Therefore, our results are not directly comparable to traditional gaze following studies.
Another promising way to assess validity is to correlate the social-cognitive ability in question to concepts that are thought to be theoretically related.
For example, it has been postulated that social-cognitive abilities are predicted by family-level variables and children's language abilities.
In our sample, caregivers were asked to fill out a short questionnaire on children's daily surroundings (e.g., sibling composition; information on childcare).
In addition, a subset of our remote child sample completed a receptive vocabulary task [@bohnorev]. We will report the association between these family-level factors and children's language skills on their gaze understanding abilities in a separate paper (in preparation). In order to match our fine-grained, parametric measure of gaze understanding, however, we hope that future technological advances will enable the continuous measures of children's real-life social surroundings [@barr2020screen; @long2020detecting; @rogoff2018importance].

@@

We have presented a new experimental paradigm to study gaze understanding across the lifespan.
Our gaze understanding task captures individual differences and shows highly satisfactory psychometric properties with respect to internal consistency and retest reliability.
The web-based testing infrastructure allows for standardized, portable data collection at scale, both remotely as well as in person.
Ultimately, this work shows a promising way forward toward more precise measures of cognitive development.
The data sets and the analysis code are freely available in the associated online repository (<https://github.com/ccp-eva/gazecues-methods>).
A demo version of the task is available at the following website: <https://ccp-odc.eva.mpg.de/gafo-demo/>.
The implementation architecture (JavaScript and HTML code) and the materials can be accessed in the following repository: <https://github.com/ccp-eva/gafo-demo>.
These resources allow interested researchers to use, extend and adapt the task.

>>>>

---
title             : "Measuring individual differences in the understanding of gaze cues across the lifespan [Julia: ich würde mehr info in den titel packen - sowas wie reliable, open-source, web-based task to measure individual differences in gaze cue understanding in children (vll sogar age range) and adults. Ich denke bei do nem methods paper ist es wichtig das man weiß was phase ist basierend auf dem Titel. Vll willst DU auch ein Acronym? :)]"
shorttitle        : "gaze understanding"
author:

@@

  - name          : "Daniel M. Haun"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"
authornote: |
abstract: |
  Traditional measures of social cognition used in developmental research often lack satisfactory psychometric properties and are not designed to capture variation *between* individuals [Julia: hab die zwei Sätze raus gelöscht weil nicht unbedingt relevant für methods paper]. Here, we present a new task to study individual differences in social cognition in children and adults. We concentrate on an essential ability that is involved in many social-cognitive reasoning processes: gaze understanding – the ability to locate and use the attentional focus of an agent. Our interactive browser-based task works across devices and enables in-person and remote testing. The implemented spatial layout allows for discrete and continuous measures of participants’ click imprecision and is easily adaptable to different study requirements. 
   Our task induces inter-individual differences in a child (N = 387) and an adult (N = 236) sample. Our two study versions and data collection modes yield comparable results that show substantial developmental gains: the older children are, the more accurately they locate the target. High internal consistency and test-retest reliability estimates underline that the captured variation is systematic. This work shows a promising way forward in the study of individual differences in social cognition and will help us explore the in(ter)dependence of our core social-cognitive processes in greater detail.
  All materials and code are freely accessible to interested researchers. A version of the task is accessible online: https://ccp-odc.eva.mpg.de/gafo-demo/.

@@

Designing this task required a new testing infrastructure.
We designed the task as an interactive browser-based web app [Julia: hier noch folgendes paper zitieren: http://dx.doi.org/https://doi.org/10.1080/15248372.2015.1061528; falls du den ansatz noch mehr "legitimieren" willst, kannst du auch noch eins oder zwei meiner tablet paper zitieren, zB: https://onlinelibrary.wiley.com/doi/10.1111/desc.13049  oder https://www.nature.com/articles/s41562-021-01145-1. Zeigt einfach dass das schon "erfolgreich" eingesetzt wird.].
This greatly increased the flexibility with which we could modify the stimuli on a trial-by-trial basis.
Furthermore, because the task is largely self-contained, it is much more controlled and standardized.
Most importantly, it makes the task portable: testing is possible in-person using tablets but also remotely via the internet (no installation needed).
As such, it provides a solid basis to study individual differences in gaze understanding across ages at scale.
We make the task and its source code openly accessible for other researchers to use and modify.

@@

[Julia: Kannst du in Figure 2 panel F noch chance linie mit rein machen?]

@@

Adding complexity to our gaze understanding task might increase individual differences in an adult sample.
Researchers could consider adjusting the task difficulty, for example, by changing the target speed or implementing a saccade version (i.e., where the agent only briefly looks at the target position and then back to the center).
Another task extension could realize differential feedback dependent on the participant's accuracy.
By inducing more between-person variance, reliability estimates could further increase [Julia: hier vll nochmal sagen, dass die momentanen properties abers chon gut sind].

As we could see in the split-half reliability calculations, the more accurately the statistical method represents the task structure, the higher our reliability estimates are.
Therefore, we argue that future research should aim at implementing statistical analyses that mirror the experimental design in its complexity [Julia: Hier noch das zitieren: https://psyarxiv.com/xr7y3/].
Computational cognitive models are a promising approach forward.
We could take advantage of all available information and model variation between and within individuals in an even more fine-grained and psychologically interpretable manner.
Computational frameworks could also be used to model performance and their underlying cognitive processes across tasks.
With nested hierarchical models, we could assess the systematic relation between various social-cognitive abilities and recover potentially shared structures between cognitive processes [@bohn2022individual].

[Julia: Hier oder weiter oben eingebaut könntest du nochmal die wishlist aufgreifen und sagen in wie fern der task die Forderungen darin abdeckt - muss nicht sein, aber wäre denke ich nett weil das so ein coole Rechfertigung ist.]

In addition to the new task design itself, we designed a new testing infrastructure.
Our gaze understanding task is presented as an interactive web app.
This enables presentation across devices without any prior installation.
Stimuli presentation is achieved through the use of SVGs.
This has several advantages: the aspect ratio and stimulus quality are kept constant no matter which size the web browser displays.
The cartoon-like presentation makes the task engaging for children and adults alike.
Most importantly, we can dynamically modify the stimulus details (e.g., target positions) on a trial-by-trial basis.
Presented agents, voice-over instructions and objects can be easily adapted for future task modifications or specific linguistic and cultural settings.

@@

After having probed our new testing infrastructure and psychometric properties, the next step will be to ensure the validity of our gaze understanding task.
Validity is often assessed by looking at concurrent relations between measures.
Studies on gaze following traditionally present children with a simultaneous, congruent movement of eyes and head orientation.
In our study, however, children only observe a subtle eye movement.
Therefore, our results are not directly comparable to traditional gaze following studies.
Another promising way to assess validity is to correlate the social-cognitive ability in question to concepts that are thought to be theoretically related.
For example, it has been postulated that social-cognitive abilities are predicted by family-level variables and children's language abilities [Julia: hier noch eine ziation jeweils, also einmal family einmal language].
[Julia: Hab das alles raus genommen - macht meiner meinung nach wenig sinn das an zu teasern und dann zu sagen das kommt in nem anderen paper. ich als reviewer oder editor würde sagen: pack das hie rein. Daher eher sagen: future research should probe these realtions (family level und language)].

@@

We have presented a new experimental paradigm to study gaze understanding across the lifespan.
Our gaze understanding task captures individual differences and shows highly satisfactory psychometric properties with respect to internal consistency and retest reliability.
The web-based testing infrastructure allows for standardized, portable data collection at scale, both remotely as well as in person.
Ultimately, this work shows a promising way forward toward more precise measures of cognitive development.
The data sets and the analysis code are freely available in the associated online repository (<https://github.com/ccp-eva/gazecues-methods>).
A demo version of the task is available at the following website: <https://ccp-odc.eva.mpg.de/gafo-demo/>.
The implementation architecture (JavaScript and HTML code) and the materials can be accessed in the following repository: <https://github.com/ccp-eva/gafo-demo>.
These resources allow interested researchers to use, extend and adapt the task. [Julia: Du könntest überlegen hier oder weiter oben noch einen screenshot von der demo version als figure zu nehmen. Das macht die usability nochmal konkreter.]
